Weighted Model Counting Without Parameter Variables

### 2021-05-13 doc/paper3/paper.tex
Remove data

\begin{lemma} \label{lemma:sum}
  For any pseudo-Boolean function $f\colon 2^X \to \mathbb{R}$, we have that
  $(\exists_X f)(\emptyset) = \sum_{Y \subseteq X} f(Y)$.
\end{lemma}
\begin{proof}
  If $X = \{x\}$, then
  \[
    (\exists_xf)(\emptyset) = (f|_{x=1} + f|_{x=0})(\emptyset) =
    f|_{x=1}(\emptyset) + f|_{x=0}(\emptyset) = \sum_{Y \subseteq \{x\}} f(Y).
  \]
  This easily extends to $|X| > 1$ by the definition of projection on sets of
  variables.
\end{proof}

@@

\textsf{DPMC} is run with tree decomposition-based planning and ADD-based
execution---the best-performing combination in the original set of experiments
\cite{DBLP:conf/cp/DudekPV20}. We use a single iteration of \textsf{htd}
\cite{DBLP:conf/cpaior/AbseherMW17} to generate approximately optimal tree
decompositions---we found that this configuration is efficient enough to handle
huge instances, and yet the width of the returned decomposition is unlikely to
differ from optimal by more than one or two. We also enabled \textsf{DPMC}'s
greedy mode. This mode (which was not part of the original paper
\cite{DBLP:conf/cp/DudekPV20}) optimises the order in which pseudo-Boolean
functions are multiplied by prioritising functions with small representations.

@@

Overall, transforming WMC instances to the PBP format allows us to significantly
simplify each instance. This transformation is particularly effective on
\texttt{bklm16}, allowing it to surpass \texttt{cd06} and become the new state
of the art. While there is a similarly significant reduction in the number of
variables for \texttt{d02}, the performance of $\textsf{DPMC}+\texttt{d02}$ is
virtually unaffected. It is also worth noting that there was no observable
difference in the width of the project-join tree used by DPMC
\cite{DBLP:conf/cp/DudekPV20} before and after applying
\cref{alg:transformation}---the observed performance improvement is more likely
related to the variable ordering heuristic used by ADDs. Finally, while our
transformation makes it possible to use \texttt{cd05} and \texttt{cd06} with
\textsf{DPMC}, the two combinations remain inefficient.

@@

\subsubsection{Acknowledgments.}
The first author was supported by the EPSRC Centre for Doctoral Training in
Robotics and Autonomous Systems, funded by the UK Engineering and Physical
Sciences Research Council (grant EP/L016834/1). The second author was supported
by a Royal Society University Research Fellowship. This work has made use of the
resources provided by the Edinburgh Compute and Data Facility (ECDF)
(\url{http://www.ecdf.ed.ac.uk/}).

>>>>

\begin{lemma} \label{lemma:sum}
  For any pseudo-Boolean function $f\colon 2^X \to \mathbb{R}$, we have that
  $(\exists_X f)(\emptyset) = \sum_{Y \subseteq X} f(Y)$.
\end{lemma}
\begin{proof}
  If $X = \{x\}$, then $(\exists_xf)(\emptyset) = (f|_{x=1} +
  f|_{x=0})(\emptyset) = f|_{x=1}(\emptyset) + f|_{x=0}(\emptyset) = \sum_{Y
    \subseteq \{x\}} f(Y)$. This easily extends to $|X| > 1$ by the definition
  of projection on sets of variables.
\end{proof}

@@

\textsf{DPMC} is run with tree decomposition-based planning and ADD-based
execution---the best-performing combination in the original set of experiments
\cite{DBLP:conf/cp/DudekPV20}. We use a single iteration of \textsf{htd}
\cite{DBLP:conf/cpaior/AbseherMW17} to generate approximately optimal tree
decompositions---we found that this configuration is efficient enough to handle
huge instances, and yet the width of the returned decomposition is unlikely to
differ from optimal by more than one or two. We also enabled \textsf{DPMC}'s
greedy mode. This mode (which was not part of the original paper
\cite{DBLP:conf/cp/DudekPV20}) optimises the order in which ADDs are multiplied
by prioritising those with small representations.

@@

It is also worth noting that there was no observable difference in the width of
the project-join tree used by DPMC \cite{DBLP:conf/cp/DudekPV20} (which is
equivalent to the treewidth of the primal/Gaifman graph of the input formula)
before and after applying \cref{alg:transformation}---the observed performance
improvement is more likely related to the variable ordering heuristic used by
ADDs.\footnote{The data on this (along with the implementation of
  \cref{alg:transformation}) is available at
  \url{https://github.com/dilkas/wmc-without-parameters}.}

Overall, transforming WMC instances to the PBP format allows us to significantly
simplify each instance. This transformation is particularly effective on
\texttt{bklm16}, allowing it to surpass \texttt{cd06} and become the new state
of the art. While there is a similarly significant reduction in the number of
variables for \texttt{d02}, the performance of $\textsf{DPMC}+\texttt{d02}$ is
virtually unaffected. Finally, while our
transformation makes it possible to use \texttt{cd05} and \texttt{cd06} with
\textsf{DPMC}, the two combinations remain inefficient.

@@

\subsubsection{Acknowledgments.}
We thank the anonymous reviewers for their helpful comments. The first author
was supported by the EPSRC Centre for Doctoral Training in Robotics and
Autonomous Systems, funded by the UK Engineering and Physical Sciences Research
Council (grant EP/L016834/1). The second author was supported by a Royal Society
University Research Fellowship. This work has made use of the resources provided
by the Edinburgh Compute and Data Facility (ECDF)
(\url{http://www.ecdf.ed.ac.uk/}).
